---
title: "p8105_hw1_rw2708"
author: "Ran Wang"
date: "9/13/2019"
output: github_document
---

# Problem 1
The chunk below creates a dataframe containing a sample of size 8 from a 
random normal variable, a logical vector indicating whether elements of the sample are greater than 0, a character vector of length 8, and a factor vector of length 8, with 3 different factor “levels”. Besides, I also computed the mean of each variable in this dataframe.

```{r Problem1_df}
library(tidyverse)

hw1_df = tibble(
  norm_samp = rnorm(8),
  vec_logical= norm_samp > 0,
 vec_char= c("1","2","3","4","5","6","7","8"),
  vec_factor=factor(c("high","medium","low","high","medium","low","high","medium"))
)


mean_samp = mean(pull(hw1_df, norm_samp))
mean_samp = mean(pull(hw1_df, vec_logical))
mean_samp = mean(pull(hw1_df, vec_char))
mean_samp = mean(pull(hw1_df, vec_factor))
```


We can get the mean of numerical and logical variables, but we can not calculate the mean for character and factor variables.

Here is a code chunk that applies the as.numeric function to the logical, character, and factor variables.

```{r converting1}

as.numeric(pull(hw1_df, vec_logical))
as.numeric(pull(hw1_df, vec_char))
as.numeric(pull(hw1_df, vec_factor))

```

Explanation for the above code chunk: logical variables, character variables, and factor variables are all converted to numeric ones. Reason: the as.numeric function serves to convert these variables to numeric form. This explains what happened when I try to take the mean, since it is only reasonable to compute the mean of numeric variables, instead of the character and factor variables.

The second code chunk is associated with the conversion of logical vectors.

```{r converting2}
numeric_convert <- as.numeric(pull(hw1_df, vec_logical))
numeric_multiply=numeric_convert*pull(hw1_df,norm_samp)

factor_convert <- as.factor(pull(hw1_df, vec_logical))
factor_multiply=factor_convert*pull(hw1_df,norm_samp)

factor_convert <- as.factor(pull(hw1_df, vec_logical))
factor_numeric <- as.numeric(factor_convert)
factor_numeric_multiply=factor_numeric*pull(hw1_df,norm_samp)

```

# Problem 2

*First I create a dataframe containing variables for my plots.

```{r problem2_df}
hw1_problem2_df = tibble(
  x= rnorm(500),
  y= rnorm(500),
 vec_logical= x+y>1,
 vec_numeric= as.numeric(vec_logical),
 vec_factor= as.factor(vec_logical)
 )
```
 
Here is a short descriptive of my vector.
  
*The dataset has `r nrow(hw1_problem2_df)` and `r ncol(hw1_problem2_df)`

*The mean of x is `r mean(pull(hw1_problem2_df,x))`
*The median of x is `r mean(pull(hw1_problem2_df,x))`
*The standard deviation of x is `r mean(pull(hw1_problem2_df,x))`
*The proportion of cases for which x + y > 1 is `mean(pull(hw1_problem2_df,vec_logical))`

The first is a scatterplot of `y` vs `x` (color points using logical variables).
```{r yx_scatter1}
ggplot(hw1_problem2_df, aes(x = x, y = y,color=vec_logical)) + geom_point()
ggsave("scatter_plot.pdf", height = 4, width = 6)
```

The second is a scatterplot of `y` vs `x` (color points using numeric variables).
```{r yx_scatter2}
ggplot(hw1_problem2_df, aes(x = x, y = y,color=vec_numeric)) + geom_point()
```

The third scatterplot of `y` vs `x` (color points using the factor variables).
```{r yx_scatter3}
ggplot(hw1_problem2_df, aes(x = x, y = y,color=vec_factor)) + geom_point()
```


