---
title: "p8105_hw1_rw2708"
author: "Ran Wang"
date: "9/13/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

set.seed(123)
```

# Problem 1
The chunk below creates a dataframe containing a sample of size 8 from a 
random normal variable, a logical vector indicating whether elements of the sample are greater than 0, a character vector of length 8, and a factor vector of length 8, with 3 different factor “levels”. Besides, I also computed the mean of each variable in this dataframe.



```{r Probelm_1, collapse = TRUE}
hw1_df = tibble(
  norm_samp = rnorm(8,0,1),
  vec_logical = norm_samp > 0,
  vec_char = c("1","2","3","4","5","6","7","8"),
  vec_factor = factor(c("high","medium","low","high","medium","low","high","medium"))
)

hw1_df

mean_norm = mean(pull(hw1_df, norm_samp))
mean_norm

mean_logical = mean(pull(hw1_df, vec_logical))
mean_logical

mean_char = mean(pull(hw1_df, vec_char))
mean_char 

mean_factor = mean(pull(hw1_df, vec_factor))
mean_factor 
```


We can get the mean of numerical and logical variables, but we can not compute the mean for character and factor variables.

Here is a code chunk that applies the as.numeric function to the logical, character, and factor variables.

```{r converting1, eval=FALSE}

as.numeric(pull(hw1_df, vec_logical))
as.numeric(pull(hw1_df, vec_char))
as.numeric(pull(hw1_df, vec_factor))

```

Explanation for the above code chunk: logical variables, character variables, and factor variables are all converted to numeric ones. Reason: the as.numeric function serves to convert these variables to numeric form. This explains what happened when I try to take the mean, since it is only reasonable to compute the mean of numeric variables, instead of the character and factor variables. After the conversion, I can calculate the mean of all of them. 

The second code chunk is associated with the conversion of logical vectors.

```{r converting2, warning = FALSE, , collapse = TRUE}
numeric_convert <- as.numeric(pull(hw1_df, vec_logical))
numeric_multiply = numeric_convert * pull(hw1_df,norm_samp)
numeric_multiply

factor_convert <- as.factor(pull(hw1_df, vec_logical))
factor_multiply = factor_convert * pull(hw1_df,norm_samp)
factor_multiply   #factor cannot be multiplied by numerical variables


factor_convert <- as.factor(pull(hw1_df, vec_logical))
factor_numeric <- as.numeric(factor_convert)
factor_numeric_multiply = factor_numeric * pull(hw1_df,norm_samp)
factor_numeric_multiply 

```

# Problem 2

*First I create a dataframe containing variables for my plots.

```{r problem2_df}
hw1_problem2_df = tibble(
  x = rnorm(500,0,1),
  y = rnorm(500,0,1),
 vec_logical = (x + y) > 1,
 vec_numeric = as.numeric(vec_logical),
 vec_factor = as.factor(vec_logical)
 )
```

Here is a short descriptive of my vector.
  
* The dataset has `r nrow(hw1_problem2_df)` rows and `r ncol(hw1_problem2_df)` columns.

* The mean of x is `r mean(pull(hw1_problem2_df,x))`

* The median of x is `r median(pull(hw1_problem2_df,x))`

* The standard deviation of x is `r sd(pull(hw1_problem2_df,x))`

* The proportion of cases for which x + y > 1 is `r mean(pull(hw1_problem2_df,vec_logical))`


The first is a scatterplot of `y` vs `x` (color points using logical variables).

```{r yx_scatter1}
ggplot(hw1_problem2_df, aes(x = x, y = y,color = vec_logical)) + geom_point()
ggsave("scatter_plot.pdf", height = 4, width = 6)
```


The second is a scatterplot of `y` vs `x` (color points using numeric variables).
```{r yx_scatter2}
ggplot(hw1_problem2_df, aes(x = x, y = y,color = vec_numeric)) + geom_point()
```


The third scatterplot of `y` vs `x` (color points using the factor variables).
```{r yx_scatter3}
ggplot(hw1_problem2_df, aes(x = x, y = y,color = vec_factor)) + geom_point()
```

